---
title: "Tidy Tuesday"
output: 
  html_document:
    toc: FALSE
---

```{r, warning = FALSE, message = FALSE, results = FALSE}
library("tidytuesdayR")
library(tidyverse)
library(dplyr)
library(ggplot2)
```
Load the data
```{r, warning = FALSE, message = FALSE, results = FALSE}
papers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/papers.csv')
authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/authors.csv')
programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/programs.csv')
paper_authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_authors.csv')
paper_programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_programs.csv')
```

Examine first few rows of data
```{r, , warning = FALSE, message = FALSE, results = FALSE}
head(authors)
head(papers)
head(programs)
head(paper_authors)
head(paper_programs)
```

Merge datasets
```{r, warning = FALSE, message = FALSE, results = FALSE}
colnames(papers) #key = paper 
colnames(paper_authors) #key = paper
colnames(authors) #key = author
colnames(paper_programs) #key = paper
colnames(programs) #key = program

joined_df <- left_join(papers, paper_authors) %>% 
  left_join(authors) %>% 
  left_join(paper_programs) %>% 
  left_join(programs)
```

This form is tidy
```{r}
summary(joined_df)
tail(joined_df)
```

Examine program variables: program, description of program, program category
```{r}
table(joined_df$program)
table(joined_df$program_desc)
table(joined_df$program_category, useNA = "ifany") #should be a factor

```

Cleaning and wrangling
```{r}
cleaned_data <- joined_df %>%
  mutate(program_category = as.factor(program_category))
```

What is the trend of working papers over time? Seems like increasing every year, with a pretty even breakdown by month. Could do stat test.
```{r}
NumPapers <- cleaned_data %>%
  group_by(year, month) %>%
  summarise(NumberofPapers = n_distinct(paper)) #to avoid double counting same paper when two authors

ggplot(data=NumPapers, aes(x=year, y=NumberofPapers, fill = as.factor(month))) +
  geom_bar(stat="identity") 
```

Who has the most working papers total? Jonathan Gruber
```{r}
cleaned_data %>%
  group_by(name) %>%
  summarise(NumberofPapers = n()) %>%
  arrange(desc(NumberofPapers)) 
```
Who has the most working papers published in one month? Janet M Currie
```{r, warning = FALSE, message = FALSE}
cleaned_data %>%
  group_by(year, month, name) %>%
  summarise(NumberofPapers = n()) %>%
  arrange(desc(NumberofPapers)) 
```
What kind of working papers are being published by the top 5 authors? Mostly microeconomics.
```{r, warning = FALSE, message = FALSE}
TopAuthors <- cleaned_data %>%
  group_by(program_category) %>%
  filter(name %in% c("Jonathan Gruber", "James J Heckman", "Daron Acemoglu", "Janet M Currie", "Michael D Bordo")) %>%
  summarise(NumberofPapers = n())

ggplot(data=TopAuthors, aes(x=program_category, y=NumberofPapers)) +
  geom_bar(stat="identity") +
  ggtitle("Program Categories of the Working Papers Published by the Top 5 Authors")
```



